---
title: "LANDFIRE R Demos"
author: "Randy Swaty and Megan Dettenmaier"
---

We are excited to share some R code bits that we think will help you in your use of LANDFIRE data to assess and visualize ecological patterns on your landscape  Over time we will be adding pages that present code we have developed over time for several LANDFIRE datasets including:

1. [Biophysical Settings](https://landfire.gov/vegetation/bps)
2. [Existing Vegetation Type](https://landfire.gov/vegetation/evt)
3. [Existing Vegetation Cover](https://landfire.gov/vegetation/evc)
4. [Existing Vegetation Height](https://landfire.gov/vegetation/evh)
5. [Fuels](https://landfire.gov/fuel)
6. [Annual Disturbance](https://landfire.gov/disturbance/annualdisturbance)

We share these bits of code in hopes of not only helping you accelerate your use of LANDFIRE data, but also as a way to start a conversation.  We'd love to hear your suggestions, and stories of LANDFIRE data processing with R!






<br>

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}


library(sf)
library(terra)


```


```{r read shapefile, message=FALSE, warning=FALSE, include=FALSE}
#  read shape
shp <- st_read("inputs/map_zone_35.shp") %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()


```

```{r locator map, message=FALSE, warning=FALSE, echo=FALSE}
library(tmap)

# toggle tmap mode to interactive viewing
tmap_mode("view")

 quickmap <- qtm(shp, 
                 col  = "darkgreen", 
                 fill = NULL, 
                 check.and.fix = TRUE, 
                 title = 'LANDFIRE Map Zone 35')

 quickmap






```

## How to use this web report

* All maps, input datasets and further assistance can be obtained by contacting the author, [Randy Swaty](mailto:rswaty@tnc.org){target="blank"}.
* Review is ongoing.
* To share this web app, simply share the hyperlink.
* Toggle between dark and light display of this site in the upper right corner of the page.
* Learn more about LANDFIRE and The Nature Conservancy [here](about.qmd){target="blank}. 
