---
title: "LANDFIRE R Demos"
author: "Randy Swaty and Megan Dettenmaier"
---

<font size="4"> 
**These LANDFIRE-Powered Landscape Assessments allow users to start exploring trends, landscape conversion and natural resource management options. **

</font> 


## Goals of this demonstration


## Demonstration Landscape



<br>

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}


library(sf)
library(terra)


```


```{r read shapefile, message=FALSE, warning=FALSE, include=FALSE}
#  read shape
shp <- st_read("inputs/map_zone_35.shp") %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()


```

```{r locator map, message=FALSE, warning=FALSE, echo=FALSE}
library(tmap)

# toggle tmap mode to interactive viewing
tmap_mode("view")

 quickmap <- qtm(shp, 
                 col  = "darkgreen", 
                 fill = NULL, 
                 check.and.fix = TRUE, 
                 title = 'LANDFIRE Map Zone 35')

 quickmap






```

## How to use this web report

* All maps, input datasets and further assistance can be obtained by contacting the author, [Randy Swaty](mailto:rswaty@tnc.org){target="blank"}.
* Review is ongoing.
* To share this web app, simply share the hyperlink.
* Toggle between dark and light display of this site in the upper right corner of the page.
* Learn more about LANDFIRE and The Nature Conservancy [here](about.qmd){target="blank}. 
