---
title: "Make hexagon maps"
---

## Skills learned

Use this page to:

- Create hexagonal grids to summarize dominant LANFDIRE vegetation types across example landscape, enabling spatial analysis beyond raster cell comparisons

- Build an interactive Leaflet map with toggle display to visualize dominant BpS and EVT within each hexagon


## Set up

For this work we will use the Area of Interest shapefile, and pre-downloaded BpS and EVT rasters.  Before you can complete this page you will need to complete the work on the [Download data](docs/download.html){target="blank"} and [Prepare rasters](docs/prep_rasters.html){target="blank"} pages.

### Code to load packages and inputs

```{r}
#| label: Set up
#| message: false
#| warning: false
#| echo: true
#| results: "hide"
#| code-overflow: wrap

# Load packages
library(exactextractr)
library(jsonlite)
library(leaflet)
library(leaflet.providers)
library(RColorBrewer)
library(sf)
library(terra)
library(tidyverse)


# read in Area of Interest (aoi) shapefile

shp <- st_read("inputs/map_zone_35.shp", quiet = TRUE) %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()

# read in .csv of CONUS-wide attributes for later joining to AoI BpS table

bps_conus_atts <- read_csv("inputs/LF20_BPS_220.csv")

# another way is directly from landfire.gov
# bps_url <- "https://landfire.gov/sites/default/files/CSV/LF2016/LF16_BPS.csv" # will likely get warning, but it's OK
# bps_conus_atts <- read.csv(bps_url)


evt_conus_atts <- read_csv("inputs/LF23_EVT_240.csv")

# evt_url <- "https://landfire.gov/sites/default/files/CSV/2024/LF2024_EVT.csv" # will likely get warning, but it's OK
# evt_conus_atts <- read.csv(evt_url)

# load in landfire stacked raster
stacked_rasters <- rast("inputs/landfire_data.tif")

# "split" cropped and masked stacked raster into separate layers
for(lyr in names(stacked_rasters)) assign(lyr, stacked_rasters[[lyr]])





```


### Make hexgrid ~ 25,000 acres a piece

In this step we make a hexgrid map for our area of interest and confirm hexagon size. 

```{r}
#| label: Make hexgrid ~ 25,000 acres a piece  # This sets a label or title for the code chunk, describing its purpose—creating a hexagonal grid with cells approximately 25,000 acres in size.
#| echo: true       # Ensures that the code itself will be displayed in the output.
#| message: false   # Suppresses messages generated during code execution (e.g., package loading messages).
#| warning: false   # Suppresses warning messages from being shown in the output.
#| code-overflow: wrap  # Wraps long lines of code instead of overflowing, improving readability.
#| include: true    # Includes both the code and its output in the final rendered document.
#| results: hide    # Hides the output of the code chunk, even though the code is included.
#| eval: true       # Evaluates (runs) the code when the document is rendered.



# Hexagon side length in meters for ~25,000 ac hexagons
hex_side <- 10930

# Create full hex grid
hex_grid <- st_make_grid(shp,
                         cellsize = hex_side,
                         square = FALSE,
                         what = "polygons") %>%
                          st_sf()

hex_grid_clipped <- st_intersection(hex_grid, shp) %>%
  mutate(index = row_number())


plot(st_geometry(hex_grid_clipped))


# Calculate area in square meters
hex_grid_size <- hex_grid %>%
  mutate(area_m2 = st_area(.),
         area_acres = as.numeric(area_m2) / 4046.86)


```


### Calculate majority type per hexagon

Below we use the extactextractr package to extract the majority (most dominant) BpS value and EVT value per hexagon, the we join in a minimal number of attributes (e.g., EVT_NAME) for mapping demo. 

```{r}
#| label: extract bps and evt info per hexagon
#| echo: true       # Ensures that the code itself will be displayed in the output.
#| message: false   # Suppresses messages generated during code execution (e.g., package loading messages).
#| warning: false   # Suppresses warning messages from being shown in the output.
#| code-overflow: wrap  # Wraps long lines of code instead of overflowing, improving readability.
#| include: true   # Includes both the code and its output in the final rendered document.
#| results: hide    # Hides the output of the code chunk, even though the code is included.
#| eval: false      # Evaluates (runs) the code when the document is rendered.




# bps
bps_majority_hex <- exact_extract(US_200BPS, hex_grid_clipped, 'majority', append_cols = "index") %>%
  left_join(select(bps_conus_atts,
                   VALUE,
                   BPS_MODEL,
                   BPS_NAME,
                   FRI_ALLFIR),
            by = c('majority' = 'VALUE'))

# evt
evt_majority_hex <- exact_extract(US_240EVT, hex_grid_clipped, 'majority', append_cols = "index") %>%
  left_join(select(evt_conus_atts,
                   VALUE,
                   EVT_NAME,
                   EVT_PHYS),
            by = c('majority' = 'VALUE'))

# Join both BpS and EVT attributes to hex shapefile
hexs_bps_evt <- hex_grid_clipped %>%
  left_join(bps_majority_hex, by = 'index') %>%
  left_join(evt_majority_hex, by = 'index')


# save the shapefile for mapping in non-R applications or to read back into R

st_write(hexs_bps_evt, "outputs/bps_evt_hexs.shp")


```

<br>

## Make the map

Below we use our attributed hexgrid to make a clickable leaflet map with the BpS and EVT layers for our area of interest.

```{r}
#| label: bps and evt map
#| echo: true       # Ensures that the code itself will be displayed in the output.
#| message: false   # Suppresses messages generated during code execution (e.g., package loading messages).
#| warning: false   # Suppresses warning messages from being shown in the output.
#| code-overflow: wrap  # Wraps long lines of code instead of overflowing, improving readability.
#| include: true   # Includes both the code and its output in the final rendered document.
#| eval: true       # Evaluates (runs) the code when the document is rendered.
#| fig-width: 12
#| fig-height: 11


hexs_bps_evt <- st_read("outputs/bps_evt_hexs.shp") 

hexs_bps_evt <- st_transform(hexs_bps_evt , crs = 4326)

# Create a dynamic color palette
bps_names <- unique(hexs_bps_evt$BPS_NAME)
evt_names <- unique(hexs_bps_evt$EVT_NAME)

# Use a colorblind-friendly base palette and expand it
base_palette <- RColorBrewer::brewer.pal(8, "Set2")  # Set2 is colorblind-friendly
bps_colors <- colorRampPalette(base_palette)(length(bps_names))

pal_bps <- colorFactor(
  palette = bps_colors,
  domain = bps_names
)

# Repeat for EVT
evt_colors <- colorRampPalette(base_palette)(length(evt_names))

pal_evt <- colorFactor(
  palette = evt_colors,
  domain = evt_names
)




# Get the bounding box of the shapefile
bbox <- st_bbox(hexs_bps_evt)

# Convert bounding box to named list
bbox_list <- as.list(bbox)

# Convert named list to JSON
json_output <- toJSON(bbox_list, keep_vec_names = TRUE)



leaflet(hexs_bps_evt) %>%
  addTiles() %>%

  fitBounds(lng1 = bbox_list$xmin, lat1 = bbox_list$ymin, lng2 = bbox_list$xmax, lat2 = bbox_list$ymax) %>%

  addPolygons(
    fillColor = ~pal_bps(BPS_NAME),
    color = "#BDBDC3",
    weight = 1,
    opacity = 1,
    fillOpacity = 1.0,
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#666",
      fillOpacity = 1.0,
      bringToFront = TRUE
    ),
    label = ~BPS_NAME,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    ),
    group = "Biophysical Settings"
  ) %>%
  addPolygons(
    fillColor = ~pal_evt(EVT_NAME),
    color = "#BDBDC3",
    weight = 1,
    opacity = 1,
    fillOpacity = 1.0,
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#666",
      fillOpacity = 1.0,
      bringToFront = TRUE
    ),
    label = ~EVT_NAME,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto"
    ),
    group = "Existing Vegetation Types"
  ) %>%
addLayersControl(
    overlayGroups = c("Biophysical Settings", "Existing Vegetation Types"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup("Existing Vegetation Types") %>% # Hide this group initially
  addScaleBar(position = "bottomleft", options = scaleBarOptions(metric = TRUE, imperial = FALSE))



```



