---
title: "Make hexagon maps"
---

## Skills learned

Use this page to:

- Create hexagonal grids to summarize dominant LANFDIRE vegetation types across example landscape, enabling spatial analysis beyond raster cell comparisons

- Build an interactive Leaflet map with toggle display to visualize dominant BpS and EVT within each hexagon


## Set up

For this work we will use the Area of Interest shapefile, and pre-downloaded BpS and EVT rasters.  Before you can complete this page you will need to complete the work on the [Download data](docs/download.html){target="blank"} and [Prepare rasters](docs/prep_rasters.html){target="blank"} pages.

### Code to load packages and inputs

```{r}
#| label: Set up
#| message: false
#| warning: false
#| echo: true
#| results: "hide"
#| code-overflow: wrap

# Load packages
library(exactextractr)
library(leaflet)
library(leaflet.providers)
library(RColorBrewer)
library(sf)
library(terra)
library(tidyverse)


# read in Area of Interest (aoi) shapefile

shp <- st_read("inputs/map_zone_35.shp", quiet = TRUE) %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()

# read in .csv of CONUS-wide attributes for later joining to AoI BpS table

bps_conus_atts <- read_csv("inputs/LF20_BPS_220.csv")

# another way is directly from landfire.gov
# bps_url <- "https://landfire.gov/sites/default/files/CSV/LF2016/LF16_BPS.csv" # will likely get warning, but it's OK
# bps_conus_atts <- read.csv(bps_url)


evt_conus_atts <- read_csv("inputs/LF23_EVT_240.csv")

# evt_url <- "https://landfire.gov/sites/default/files/CSV/2024/LF2024_EVT.csv" # will likely get warning, but it's OK
# evt_conus_atts <- read.csv(evt_url)

# load in landfire stacked raster
stacked_rasters <- rast("inputs/landfire_data.tif")

# "split" cropped and masked stacked raster into separate layers
for(lyr in names(stacked_rasters)) assign(lyr, stacked_rasters[[lyr]])





```


### Make hexgrid ~ 25,000 acres a piece

```{r}
#| label: Make hexgrid ~ 25,000 acres a piece
#| echo: true
#| message: false
#| warning: false
#| code-overflow: wrap
#| include: true
#| results: hide
#| eval: true


# Hexagon side length in meters for ~25,000 ac hexagons
hex_side <- 10930

# Create full hex grid
hex_grid <- st_make_grid(shp,
                         cellsize = hex_side,
                         square = FALSE,
                         what = "polygons") %>%
                          st_sf()

hex_grid_clipped <- st_intersection(hex_grid, shp) %>%
  mutate(index = row_number())


plot(st_geometry(hex_grid_clipped))


# Calculate area in square meters
hex_grid_size <- hex_grid %>%
  mutate(area_m2 = st_area(.),
         area_acres = as.numeric(area_m2) / 4046.86)


```


### Calculate majority type per hexagon


```{r}



# bps
bps_majority_hex <- exact_extract(US_200BPS, hex_grid_clipped, 'majority', append_cols = "index") %>%
  left_join(select(bps_conus_atts,
                   VALUE,
                   BPS_MODEL,
                   BPS_NAME,
                   FRI_ALLFIR),
            by = c('majority' = 'VALUE'))

# evt
evt_majority_hex <- exact_extract(US_240EVT, hex_grid_clipped, 'majority', append_cols = "index") %>%
  left_join(select(evt_conus_atts,
                   VALUE,
                   EVT_NAME,
                   EVT_PHYS),
            by = c('majority' = 'VALUE'))

# Join both BpS and EVT attributes to hex shapefile
hexs_bps_evt <- hex_grid_clipped %>%
  left_join(bps_majority_hex, by = 'index') %>%
  left_join(evt_majority_hex, by = 'index')

```






